SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2018-09-12 scipy-1.0/paper.tex
Write background section w/ Jarrod

\author[1]{Pauli Virtanen}
\author[2,*]{Ralf Gommers}
\author[3,4]{Tyler Reddy}
\author[5]{Anne Archibald}
\author[6]{Andrew Nelson}
\author[7]{Charles Harris}
\author[8]{CJ Carey}
\author[9]{Denis Laxalde}
\author[10]{Eric Larson}
\author[11]{Eric Moore}
\author[12]{Eric Quintero}
\author[13]{Evgeni Burovski}
\author[14]{Jaime Fernández del Río}
\author[15]{Josef Perktold}
\author[16]{Josh Wilson}
\author[17]{Matthew Brett}
\author[18]{Nikolay Mayorov}
\author[19]{Warren Weckesser}
\author[20]{Matt Haberland}
\author[21]{Scott Sievert}
\author[22]{Yu Feng}
\author[23]{Antonio Horta Ribeiro}
\author[24]{Ian Henriksen}

\affil[1]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{UC Berkeley Institute for Data Science,
          Berkeley, CA, 94720, USA}
\affil[4]{Los Alamos National Laboratory,
	  Theoretical Division 6,
          Los Alamos, NM, 87545, USA}
\affil[5]{Affiliation, department, city, postcode, country}
\affil[6]{Affiliation, department, city, postcode, country}
\affil[7]{Affiliation, department, city, postcode, country}
\affil[8]{Affiliation, department, city, postcode, country}
\affil[9]{Affiliation, department, city, postcode, country}
\affil[10]{Affiliation, department, city, postcode, country}
\affil[11]{Affiliation, department, city, postcode, country}
\affil[12]{Affiliation, department, city, postcode, country}
\affil[13]{Affiliation, department, city, postcode, country}
\affil[14]{Affiliation, department, city, postcode, country}
\affil[15]{Affiliation, department, city, postcode, country}
\affil[16]{Affiliation, department, city, postcode, country}
\affil[17]{Affiliation, department, city, postcode, country}
\affil[18]{Affiliation, department, city, postcode, country}
\affil[19]{Affiliation, department, city, postcode, country}
\affil[20]{Affiliation, department, city, postcode, country}
\affil[21]{Affiliation, department, city, postcode, country}
\affil[22]{Affiliation, department, city, postcode, country}
\affil[23]{Affiliation, department, city, postcode, country}
\affil[24]{University of Texas at Austin,
           Institute for Computational Engineering and Sciences,
	   Austin, TX, 78712, USA}

@@

Python ... late 1980s

An matrix / n-dimensional array object ...
matrix.py, numeric, numarray, numpy

Paul Dubois

Travis O

Pearu

-
Gary Strangman

%https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
%

Eric and Travis V

@@

>>>>

\author[1]{Pauli Virtanen}
\author[2,*]{Ralf Gommers}
\author[3,4]{Tyler Reddy}
\author[5]{Anne Archibald}
\author[6]{Andrew Nelson}
\author[7]{Charles Harris}
\author[8]{CJ Carey}
\author[9]{Denis Laxalde}
\author[10]{Eric Larson}
\author[11]{Eric Moore}
\author[12]{Eric Quintero}
\author[13]{Evgeni Burovski}
\author[14]{Jaime Fernández del Río}
\author[15]{Josef Perktold}
\author[16]{Josh Wilson}
\author[17]{Matthew Brett}
\author[18]{Nikolay Mayorov}
\author[19]{Warren Weckesser}
\author[20]{Matt Haberland}
\author[21]{Scott Sievert}
\author[22]{Yu Feng}
\author[23]{Antonio Horta Ribeiro}
\author[24]{Ian Henriksen}
\author[3,25]{K. Jarrod Millman}
\author[3]{St\'efan J. van der Walt}

\affil[1]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{Berkeley Institute for Data Science, University of California, Berkeley, CA, 94720, USA}
\affil[4]{Los Alamos National Laboratory,
	  Theoretical Division 6,
          Los Alamos, NM, 87545, USA}
\affil[5]{Affiliation, department, city, postcode, country}
\affil[6]{Affiliation, department, city, postcode, country}
\affil[7]{Affiliation, department, city, postcode, country}
\affil[8]{Affiliation, department, city, postcode, country}
\affil[9]{Affiliation, department, city, postcode, country}
\affil[10]{Affiliation, department, city, postcode, country}
\affil[11]{Affiliation, department, city, postcode, country}
\affil[12]{Affiliation, department, city, postcode, country}
\affil[13]{Affiliation, department, city, postcode, country}
\affil[14]{Affiliation, department, city, postcode, country}
\affil[15]{Affiliation, department, city, postcode, country}
\affil[16]{Affiliation, department, city, postcode, country}
\affil[17]{Affiliation, department, city, postcode, country}
\affil[18]{Affiliation, department, city, postcode, country}
\affil[19]{Affiliation, department, city, postcode, country}
\affil[20]{Affiliation, department, city, postcode, country}
\affil[21]{Affiliation, department, city, postcode, country}
\affil[22]{Affiliation, department, city, postcode, country}
\affil[23]{Affiliation, department, city, postcode, country}
\affil[24]{University of Texas at Austin,
           Institute for Computational Engineering and Sciences,
	   Austin, TX, 78712, USA}
\affil[25]{Division of Biostatistics, University of California,
  Berkeley, CA, 94720, USA}

@@

Python is an interactive computer programming language, designed by
Guido van Rossum in the late 80s, with a heavy emphasis on readability
and rapid prototyping.
% https://legacy.python.org/search/hypermail/python-1992/0000.html
%
While it was a general purpose programming language with a ``batteries
included'' standard library, it had no special support for scientific
data structures or algorithms, unlike many of the other established computation
platforms of the time. Yet, scientists soon discovered the language's
virtues, such as its ability to wrap C and Fortran libraries and to
then drive those libraries interactively.
Scientists could thereby gain access to a wide variety of existing
computational libraries, without concerning themselves with low-level
programming concepts as as memory management.

In 1995, Jim Hugunin, a graduate student from MIT, wrote in the first
message on the Python Matrix Special Interest Group
mailing list\footnote{https://mail.python.org/pipermail/matrix-sig/1995-August/000002.html}:
\begin{quote}
There seems to be a fair amount of interest in the Python community
concerning the addition of numeric operations to Python.  My own desire is
to have as large a library of matrix based functions available as possible
(linear algebra, eigenfunctions, signal processing, statistics, etc.).  In
order to ensure that all of these libraries interoperate, there needs to
be agreement on a basic matrix object that can be used to represent arrays
of numbers.
\end{quote}
Over the next several months, conversations on that mailing
list by, among others, Jim Fulton, Jim Hugunin, Paul Dubois, Konrad
Hinsen, and Guido van Rossum, led to the creation of an array object
with arbitrarily high dimensions.  Jim Hugunin explained the utility
of Python for numerical computation\footnote{http://hugunin.net/papers/hugunin95numpy.html}:
\begin{quote}
I've used almost all of the available numerical languages at one time
or another over the past 8 years. One thing I've noticed is that over
time, the designers of these languages are steadily adding more of the
features that one would expect to find in a general-purpose
programming language.
\end{quote}
This is still one of the distinguishing features of Python for
science, and why it has been so successful in the realm of data
science: instead of adding general features to a language designed for
numerical and scientific computing, here scientific features are added
to a general purpose language instead.  This broadens the scope of
problems that can be addressed easily, the sources of data that are
readily accessible, and increases the size of the community that
develops general purpose tooling.

The availability of a standard numerical array data structure in
Python led to a sudden and rapidly growing number of scientific
packages becoming available for solving various common numeric
problems.

A number of these packages were written by graduate students and
postdoctoral researchers to solve the very practical research problems
that they faced on a daily basis.  While they had access to and were
familiar with specialized (and often commercial) systems, many found
it easier to implement the domain-specific functionality they needed
in a general purpose programming language.  And given Python's innate
ability to function as a systems language, controlling specialized or
custom-built hardware was feasible.

@@

Travis O
Pearu
-
Gary Strangman
%
%https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
%
%
Eric and Travis V
